---
title: "Assignment 3: Midterm"
---

## Dr. Walker's Lab Report

I have just ran an experiment that measures the effect of exercise on sleep habits. I was able to measure participants average hours of sleep before and after the experiment. I was also able to measure participants sleep efficiency at the end of the experiment. Since the experiment is over, I needed to identify which, if any, of the exercise methods best impacts sleep.

The packages we will be using for this assignment include: **readxl, dplyr, tibble, tidyverse, mosaic, ggplot2, supernova, and stats.**
```{r}
#| label: loading packages
library(readxl)
library(dplyr)
library(tibble)
library(tidyverse)
library(mosaic)
library(ggplot2)
library(supernova)
library(stats)
```


::: callout-important
Remember that you must install and use library for each R package in order to follow along!
:::

Here is my report.

## Data

In this section, I ran some code to load the two data sets for the analysis **below**. I am using glimpse and head to preview my two data sets to ensure they correctly have been loaded.

```{r}
#| label: Loading data sets
participant_info_midterm <- read_excel("participant_info_midterm.xlsx")
glimpse(participant_info_midterm)

sleep_data_midterm <- read_excel("sleep_data_midterm.xlsx")
glimpse(sleep_data_midterm)
head(sleep_data_midterm)
```

## Data Cleaning

In this section, we will be cleaning the data. The column names thankfully do not have to be cleaned.

In this piece of code **below**, we are standardizing the column titled Exercise_Group.

```{r}
#| label: cleaning the data by standardizing a column

participant_info_midterm<-participant_info_midterm %>%
  mutate(Exercise_Group =case_when(
    Exercise_Group=="NONE" ~"None",
    Exercise_Group=="Nonee" ~ "None",
    Exercise_Group=="N"~"None",
    Exercise_Group=="C"~"Cardio",
    Exercise_Group=="WEIGHTS"~"Weights",
    Exercise_Group=="WEIGHTZ"~"Weights",
    Exercise_Group=="WEIGHTSSS"~"Weights",
    Exercise_Group=="CW"~"Cardio+Weights",
    Exercise_Group=="C+W"~"Cardio+Weights",
    TRUE~Exercise_Group
  ))
```

In this piece of code **below**, we are standardizing the column titled Sex.

```{r}
#| label: Standaizing the sex column
participant_info_midterm<-participant_info_midterm %>%
  mutate(Sex=case_when(
    Sex=="Malee"~"Male",
    Sex=="MALE"~"Male",
    Sex=="M"~"Male",
    Sex=="Mal"~"Male",
    Sex=="Femalee"~"Female",
    Sex=="F"~"Female",
    TRUE~Sex
  ))
```

Since our data is now all clean, the last step was to merge our two data sets.

We can now see that our data is cleaned and merged **below**.

```{r}
#| label: Merging data sets
Data_Midterm<- merge(participant_info_midterm, sleep_data_midterm, by="ID")

knitr::kable(Data_Midterm)
```

## Create Derived Variables

We now want to Create a column titled Sleep_Difference with Post_Sleep - Pre_Sleep.

First we got to check if Post_Sleep and Pre_Sleep columns are numeric **below**.

```{r}
#| label: checking if columns are numeric
is.numeric(Data_Midterm$Post_Sleep) 
is.numeric(Data_Midterm$Pre_Sleep)
```

Post_Sleep IS numeric but Pre_Sleep is NOT numeric. Before we use code to turn it into numeric, we first have to clean the column further since it has letters and numbers. I did this by separating the letters and numbers into separate columns **below**.

```{r}
#| label: cleaning column before turning it numeric
Data_Midterm<- Data_Midterm %>%
  separate(
    col= Pre_Sleep,
    into= c("Pre_Sleep_Letters", "Pre_Sleep"),
    sep= "-"
  )
```

I then turned the Pre_Sleep column into numeric and deleted Pre_Sleep_Letters column. You can see here that the Pre_Sleep_Letters column is gone and we are left with Pre_Sleep Column cleaned up. I also checked if Pre_Sleep column is numeric and it is. All of this can be seen **below**.

```{r}
#| label: turning column numeric
Data_Midterm<- Data_Midterm%>% 
  mutate(across(
    c("Pre_Sleep"), 
    ~as.numeric(.)
  ))

Data_Midterm<- Data_Midterm %>% 
  select(-"Pre_Sleep_Letters")

is.numeric(Data_Midterm$Pre_Sleep)

head(Data_Midterm)
```

Now we can create our new column titled Sleep_Difference.

Furthermore, I identified that there was 14 participants with no Sleep_Difference scores. I removed them from our data leaving us with 86 participants which you can view **below**.

```{r}
#| label: removing participants with no sleep difference scores
Data_Midterm<-Data_Midterm%>%
  mutate(Sleep_Difference=Post_Sleep-Pre_Sleep)

Data_Midterm$Sleep_Difference %>% is.na() %>% which()

Data_Midterm<- Data_Midterm %>% filter(!is.na(Sleep_Difference))

knitr::kable(Data_Midterm)
```

Next we want to create a new column titled AgeGroup2 using via case_when. You can see **below** that we have successfully created this new column.

```{r}
#| label: Creating a new column for age group based on younger and older adult
Data_Midterm<-Data_Midterm%>%
  mutate(AgeGroup2=case_when(
    Age <35~"Younger_Adult",
    Age >=35~"Older_Adult"
  ))

head(Data_Midterm)
```

## Descriptive Statistics

**Below** we are exploring the descriptive statistics for Sleep_Difference.

```{r}
#| label: descriptive statistics to explore sleep difference column
Sleep_Diff_Stats<- favstats(~Sleep_Difference, data = Data_Midterm) 
Sleep_Diff_Stats
```

For Sleep_Difference column, the statistics is as follows:

-   Mean: 0.68

-   SD: 0.66

-   Min: -1.1

-   Max: 2.1

**Below** we are exploring the descriptive statistics for Sleep_Efficiency.

```{r}
#| label:  descriptive statistics to explore sleep efficiency column
Sleep_Effic_Stats<-favstats(~Sleep_Efficiency, data = Data_Midterm)
Sleep_Effic_Stats
```

For Sleep_Efficiency column, the statistics is as follows:

-   Mean: 83.78

-   SD: 5.98

-   Min: 71.7

-   Max: 101.5

**Below** we explore the pairwise means of Sleep_Difference by Exercise_Group.

```{r}
#| label: exploring pairwise means of sleep difference and exercise group columns
Sleep_Diff_by_Exercise<-favstats(Sleep_Difference~Exercise_Group, data = Data_Midterm)
Sleep_Diff_by_Exercise
```

For pairwise means, the means are as follows:

-   Cardio: 1.14

-   Cardio+Weights: 0.86

-   None: 0.048

-   Weights: 0.67

**Below** we explore the pairwise means of Sleep_Efficiency by Exercise_Group.

```{r}
#| label: exploring pariwise means of sleep efficiency and exercise group columns
Sleep_Effic_by_Exercise<-favstats(Sleep_Efficiency~Exercise_Group, data = Data_Midterm)
Sleep_Effic_by_Exercise
```

For pairwise means, the means are as follows:

-   Cardio: 85.45

-   Cardio+Weights: 86.83

-   None: 81.07

-   Weights: 81.46

## Visualizations (Plots)

I proceeded to create plots to be able to visualize things better.

```{r}
#| label: Boxplot created to look at sleep difference by exercise group
#| fig-cap: fig-Boxplot showing exercise group by sleep difference in order to look at it visually
ggplot(Data_Midterm, aes(x=Exercise_Group, y=Sleep_Difference))+
  geom_boxplot(fill="orange", color="black")+
  coord_flip() +
  labs(
    title = "Exercise Group by Sleep Difference",
    x="Exercise Type",
    y="Sleep Difference")+
theme(plot.title = element_text(size=20, family="serif", face="bold"),
             axis.title = element_text(size=15, family ="serif"),
             axis.text = element_text(size = 10, family = "serif"))
```

**Above** a box plot was created to look at sleep difference by exercise group.

```{r}
#| label: Boxplot created to look at sleep efficiency 
#| fig-cap: fig-Boxplot showing sleep efficiency by exercise group in order to look at it visually
ggplot(Data_Midterm, aes(x=Exercise_Group, y=Sleep_Efficiency))+
  geom_boxplot(fill="orange", color="black")+
  coord_flip() +
  labs(
    title = "Exercise Group by Sleep Efficiency",
    x="Exercise Type",
    y="Sleep Efficiency")+
  theme(plot.title = element_text(size=20, family="serif", face="bold"),
        axis.title = element_text(size=15, family ="serif"),
        axis.text = element_text(size = 10, family = "serif"))
```

**Above** a box plot was created to look at sleep efficiency by exercise group.

```{r}
#| label: Boxplot created to look at sleep difference by sleep efficiency 
#| fig-cap: fig-Graph with a trend line showing sleep difference by sleep efficiency in order to look at it visually
ggplot(Data_Midterm, aes(x=Sleep_Difference, y=Sleep_Efficiency))+
  geom_point(alpha=1, size=2)+
  labs(
    title = "Sleep Difference by Sleep Efficiency",
    x="Sleep Difference",
    y="Sleep Efficiency"
  )+
  geom_smooth(method = "lm", se = FALSE, color="orange")+
  theme(plot.title = element_text(size=23, family="serif", face="bold"),
        axis.title = element_text(size=15, family ="serif"),
        axis.text = element_text(size = 10, family = "serif"))
```

**Above** a graph was created with a trend line to look at sleep difference by sleep efficiency.

## T-Test's

I went ahead and conducted two t-tests.

**Below** I conducted a two sample t-test for sleep difference by sex (male vs female).

```{r}
#| label: T-test statastical method conducted  for sleep difference by sex
SleepDiffbySex_Ttest<-t.test(Sleep_Difference ~ Sex, data = Data_Midterm)
SleepDiffbySex_Ttest
```

-   Mean in Female Group: 0.78

-   Mean in Male Group: 0.55

-   *P* value: 0.12

-   The difference between males and females in their sleep differences is not significant.

**Below** I conducted a two sample t-test for sleep difference by age group (young vs older adults).

```{r}
#| label: T-Test statistical method conducted for sleep difference by age group
SleepDiffbyAge_Ttest<-t.test(Sleep_Difference ~ AgeGroup2, data = Data_Midterm)
SleepDiffbyAge_Ttest
```

-   Mean in Older Adult Group: 0.74

-   Mean in Younger Adult Group: 0.63

-   *P* value: 0.43

-   The difference between younger adult and older adult participants in their sleep difference is not significant.

## ANOVA'S

I went ahead and conducted two ANOVA's along with posthoc comparisons if the ANOVA was significant.

**Below** I conducted an ANOVA for sleep difference by exercise group and posthoc comparisons.

```{r}
#| label: ANOVA statistical method conducted for sleep difference by exercise group 
SleepDiffbyExercise_ANOVA<-aov(Sleep_Difference ~ Exercise_Group, data = Data_Midterm)

summary(SleepDiffbyExercise_ANOVA)

supernova(SleepDiffbyExercise_ANOVA)
```

Important Information for **above**:

-   F = 15.717

-   df = 85

-   *p* = .0000

-   PRE = .37

-   effect size = Small

```{r}
#| label: Tukey test conducted for posthoc for ANOVA about sleep difference by exercise group
TukeyHSD(SleepDiffbyExercise_ANOVA)
```

TukeyHSD for **above** code:

TukeyHSD:

-   No exercise is statistically different to Cardio (P=0.0000). Cardio has a greater mean difference of 1.09 than no exercise, meaning Cardio has a higher sleep difference.

-   Weights is statistically different to Cardio (P=0.03). Cardio has a greater mean difference of .47 than weights, meaning Cardio has a higher sleep difference.

-   No exercise is statistically different to Cardio + Weights (P= 0.00002). Cardio+weights has a greater mean difference of .81 than no exercise, meaning Cardio+weights has a higher sleep difference.

-   Weights is statistically different to no exercise (P=0.001). Weights has a greater mean difference of 0.62 than no exercise, meaning weights has a higher sleep difference.

-   Overall, Cardio has the best sleep difference due to having the largest mean difference (1.09) and smallest *p* value (p=0.000).

**Below** I conducted an ANOVA for sleep efficiency by exercise group and posthoc comparisons.

```{r}
#| label: ANOVA statistical method conducted for sleep efficiency by exercise group
SleepEfficbyExercise_ANOVA<-aov(Sleep_Efficiency ~ Exercise_Group, data = Data_Midterm)

summary(SleepEfficbyExercise_ANOVA)

supernova(SleepEfficbyExercise_ANOVA)

```

Important Information for **above**:

-   F = 5.93

-   df = 85

-   *p* = .001

-   PRE = .18

-   Effect size = Very Small

```{r}
#| label: Tukey test conducted ANOVA about sleep efficiency by exercise group
TukeyHSD(SleepEfficbyExercise_ANOVA)
```

TukeyHSD for **above** code:

-   No exercise is not statistically different to Cardio but trending to significance (p= 0.057). Cardio has a greater mean difference of 4.38 than no exercise, meaning Cardio has a higher sleep efficiency.

-   No exercise is statistically significant to Cardio+Weights (p=0.005). Cardio+Weights has a greater mean difference of 5.76 than no exercise, meaning Cardio+Weights has a higher sleep efficiency.

-   Weights is statistically significant to Cardio+Weights (p=0.009). Cardio+Weights has a greater mean difference of 5.38 than weights, meaning Cardio+Weights has a higher sleep efficiency.

-   Overall Cardio+Weights has the best sleep efficiency due to having a greater mean (5.76 and 5.38) and smallest p value (0.005 and 0.009)

## Synthesis & Recommendation

If I had to pick one exercise regimen to recommend to improve overall sleep, I would recommend Weights+Cardio. Regardless of sex and age, Weights+Cardio was significant in improving sleep difference before and after exercise than no exercise (P= 0.00002) and improving sleep efficiency after this exercise rather than no exercise (p=0.005). Although Cardio had a smaller p value (P=0.0000) when compared to no exercise when it comes to sleep difference before and after exercise type, it was only trending to significance for sleep efficiency (p= 0.057). Cardio did improve sleep difference before and after more than Cardio+Weights, but did not significantly improve sleep efficiency. Visually, in graph 1 and 2, you can see that Cardio+Weights are close to each other, but Cardio+Weights exceeds Cardio in sleep efficiency. It is important to consider overall sleep, and therefore, I recommend Weights+Cardio.

## Reflection

The most challenging part of this report was separating the column pre_sleep to have only the numbers for my analysis. I think I panicked, but after taking time to think and read over notes, I was able to figure it out and successfully run my code. I also felt that interpreting the pairwise comparisons was challenging as it has been a while doing posthoc, but I managed to finish! Overall, I felt confident in the rest of the task because I have been practicing my coding skills and always take my time when I do any assignment. It is like I relearn when I do assignments, which helps tremendously.
